---
title: "HW 02"
author: "Nandakumar Kuthalaraja"
format: 
  html:
    embed-resources: true
toc: true
---

```{r setup}
if (!require("pacman")) 
  install.packages("pacman")

pacman::p_load(tidyverse, ggridges, glue, scales, ggthemes, openintro, ggrepel, dsbox, janitor, fs)

# set theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))

# set width of code output
options(width = 85)

# set figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7,        # 7" width
  fig.asp = 0.618,      # the golden ratio
  fig.retina = 3,       # dpi multiplier for displaying HTML output on retina
  fig.align = "center", # center align figures
  dpi = 300             # higher dpi, sharper image
)



###All responses are in comments within the code****

```

## 1 - A new day, a new plot, a new geom

```{r}
#| label: Question 1

#read from dsbox pkg dataset
#glimpse(edibnb)

edibnb <- edibnb |>
  filter(!is.na(neighbourhood), !is.na(review_scores_rating))
median_rating <- edibnb |>
  group_by(neighbourhood) |>
  summarise(median_review_score = median(review_scores_rating, na.rm = TRUE))|>
  arrange(median_review_score)# got the median review score per neighborhood
  
#glimpse(median_rating)

neighborhood_order <- edibnb |>
  mutate(neighbourhood = factor(neighbourhood,  levels = median_rating$neighbourhood))

ggplot(neighborhood_order, aes(x = review_scores_rating, y = neighbourhood)) +
  geom_density_ridges(alpha = 0.7) +
  theme(legend.position = "none",
        plot.subtitle = element_text(size = 8)) + 
  labs(
    title = "Airbnb Review of Edinburgh Neighborhood",
    x = "Review Score (0–100)",
    y = "Neighborhood",
    subtitle = "DensityRidge Plot"
  )

```
This plot, shows the distribution of Airbnb review scores across different neighborhoods in Edinburgh. Each ridge represents a neighborhood and illustrates how review scores are spread out for listings in that area. 
The highest scoring neighborhoods tend to have a "taller" distribution curve, implying a tighter spread of high review scores compared to the bottom few.

## 2 - Foreign Connected PACs

```{r}
#| label: label-me-2

```

## 3 - Median housing prices in the US

## 4 - Expect More. Plot More.

## 5 - Mirror, mirror on the wall, who’s the ugliest of them all?
